Out-of-Sample Comparisons of Overfit Models

### 2013-12-20 paper.tex
Tweak a few sentences in the introduction

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of \oos\ comparisons when at least
one of the models is overfit---the number of its regressors
increases with the number of observations so that their ratio remains
positive.  Although overfit is sometimes used to describe the
situation where a forecaster chooses from many different models,
i.e. \textit{data-mining} or \textit{data-snooping}, we view these as
separate issues.  Procedures that account for the presence of many
models have been, and are continuing to be, developed \citep[see, for
example,][]{Whi:00,Han:05,RoW:05,HHK:10,ClM:12b}, but it is unclear
whether those procedures should themselves use in-sample or \oos\
comparisons.  Understanding the difference between in-sample and \oos\
comparisons in the context of a simple comparison between two models
is necessary before resolving any new issues that arise with multiple
comparisons.\footnote{Moreover, the empirical research that motivates
  this paper uses \textit{pseudo} \oos\ comparisons and not
  \textit{true} \oos\ comparisons.  Even if a true \oos\ comparison
  could account for some forms of data-snooping better than
  \citepos{Whi:00} \textsc{brc} or its extensions, in-sample and
  pseudo \oos\ comparisons would both be affected by the data-snooping.
  This point is also made by \citet{InK:04}.}

We focus on linear regression models for simplicity, but our basic
conclusions should be true for other models as well.  Under this
asymptotic theory, where the number of regressors $k$ increases with $T$
so that $\lim k/T$ is positive,
the \ols\ coefficient estimator is no longer
consistent or asymptotically normal \citep{Hub:73} and has positive
variance in the limit.  We show that, even so, the usual \oos\ average
is asymptotically normal and can consistently estimate the difference
between the models' generalization error, the expected loss in the
future conditional on the on the data available in period
$T$.\footnote{See, for example, \citet{HTF:08} for a discussion of
  generalization error.}  Under these asymptotics, the generalization
error does not converge to the expected performance of the pseudo-true
models, so existing in-sample and \oos\ comparisons measure different
quantities and should be expected to give different results for
reasons beyond simple size and power comparisons.  Under our limit theory,
the model that is closer to the true \dgp\ in population can forecast
worse.  In such a situation, a standard in-sample comparison would
correctly reject the null hypothesis that the benchmark is true, and
an \oos\ comparison would correctly fail to reject the null that the
benchmark is more accurate.\footnote{In a pair of papers similar to
  ours, \citet{ClM:12,ClM:12b} study in-sample and \oos\ tests that
  the larger model has nonzero coefficients that are too close to zero
  to expect the model to forecast more accurately.  Like this paper,
  they argue that the larger model can be true but less accurate.
  However, they focus on an aspect of the \dgp\ that makes this
  phenomenon likely, while we focus on the coefficient estimates that
  produce less accurate forecasts.  Moreover, the implications of our
  asymptotic theories are different and their papers do not provide
  reasons to do \oos\ comparisons.}
Also note that, in this situation, a model that performs well
in-sample can perform badly out-of-sample \emph{even if there are no
structural breaks or other forms of instability.}  Researchers
have argued that a breakdown of in-sample forecasting ability
indicates a structural break (see \citealp{BoH:99} and
\citealp{StW:03} among others), but this breakdown can be caused by overfit as
well.

@@

Finally, this paper also introduces a new method of proof for \oos\
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of \oos\ loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales when recentered, so asymptotic
results for \ned\ functions of mixing processes can be used for these
\oos\ processes with only slight modification.

>>>>

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of \oos\ comparisons when at least
one of the models is overfit---the number of its regressors
increases with the number of observations so that their ratio remains
positive.  Although overfit is sometimes used to describe the
situation where a forecaster chooses from many different models,
i.e. \textit{data-mining} or \textit{data-snooping}, we view these as
separate issues.  Procedures that account for the presence of many
models have been, and are continuing to be, developed \citep[see, for
example,][]{Whi:00,Han:05,RoW:05,HHK:10,ClM:12b}, but it is unclear
whether those procedures should themselves use in-sample or \oos\
comparisons.  Understanding the difference between in-sample and \oos\
comparisons in the context of a simple comparison between two models
is necessary before resolving any new issues that arise with multiple
comparisons. Moreover, the empirical research that motivates
this paper uses \textit{pseudo} \oos\ comparisons and not
\textit{true} \oos\ comparisons.  Even if a true \oos\ comparison
could account for some forms of data-snooping better than
\citepos{Whi:00} \textsc{brc} or its extensions, in-sample and pseudo
\oos\ comparisons would both be affected by the data-snooping, a
point also made by \citet{InK:04}.

We focus on linear regression models estimated with a fixed window
for simplicity, but our basic
conclusions should be true for other models as well.  Under this
asymptotic theory, where the number of regressors $k$ increases with $T$
so that $\lim k/T$ is positive and less than one,
the \ols\ coefficient estimator is no longer
consistent or asymptotically normal \citep{Hub:73} and has positive
variance in the limit.  We show that, even so, the usual \oos\ average
is asymptotically normal and can consistently estimate the difference
between the models' generalization error, the expected loss in the
future conditional on the on the data available in period
$T$.\footnote{See, for example, \citet{HTF:08} for a discussion of
  generalization error.}  Under these asymptotics, the generalization
error does not converge to the expected performance of the pseudo-true
models, so existing in-sample and \oos\ comparisons measure different
quantities and should be expected to give different results for
reasons beyond simple size and power comparisons.  Under our limit theory,
the model that is closer to the true \dgp\ in population can forecast
worse.  In such a situation, a standard in-sample comparison would
correctly reject the null hypothesis that the benchmark is true, and
an \oos\ comparison would correctly fail to reject the null that the
benchmark is more accurate.\footnote{In a pair of papers similar to
  ours, \citet{ClM:12,ClM:12b} study in-sample and \oos\ tests that
  the larger model has nonzero coefficients that are too close to zero
  to expect the model to forecast more accurately.  Like this paper,
  they argue that the larger model can be true but less accurate.
  However, they focus on an aspect of the \dgp\ that makes this
  phenomenon likely, while we focus on the coefficient estimates that
  produce less accurate forecasts.  Moreover, the implications of our
  asymptotic theories are different and their papers do not provide
  reasons to do \oos\ comparisons.}
Also note that, in this situation, a model that performs well
in-sample can perform badly out-of-sample even if there are no
structural breaks or other forms of instability.  Researchers
have argued that a breakdown of in-sample forecasting ability
indicates a structural break (see \citealp{BoH:99} and
\citealp{StW:03} among others), but this breakdown can be caused by overfit as
well.

@@

Finally, this paper also introduces a new method of proof for \oos\
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of \oos\ loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales when recentered, so 
the arguments used to prove asymptotic
results for \ned\ functions of mixing processes can be used for these
\oos\ processes with only slight modification.
